---
title: "What is Parliament Saying About Education?"
author: "Kaylynn Driggers"
date: "2026-04-05"
output: pdf_document
---
##Set Up
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidylog)
library(readxl)
library(socsci)
library(tidytext)
library(quanteda)
library(readr)
library(dplyr)
library(stringr)
library(irr) #for cohen's kappa
library(tinytex)
```

##Introduction

This project creates an education speech corpus from the Canadian Hansard Files from 1901-2019. 

#Part 1

## Creating a For-Loop to Create a Master Hansard Speech File
```{r, include=FALSE}
# Set the path to your master folder
master_folder <- "C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Hansard Speech Data"

# Get a list of year folders inside the master folder
folder_list <- list.dirs(master_folder, recursive = FALSE, full.names = TRUE)

# Count total number of year folders
total_years <- length(folder_list)

# Set seed for reproducibility
set.seed(2026)

# Create empty list to store sampled data
sample_list <- list()
counter <- 1

# Outer loop: go through each year folder
for (i in seq_along(folder_list)) {
  
  folder <- folder_list[i]
  year <- basename(folder)
  
  # Find all CSV files inside the year folder, including month subfolders
  file_list <- list.files(folder,
                          pattern = "\\.csv$",
                          full.names = TRUE,
                          recursive = TRUE,
                          ignore.case = TRUE)
  
  cat("Processing year", year, "(", i, "of", total_years, "):",
      length(file_list), "files found\n")
  
  # Inner loop: go through each CSV file
  for (file in file_list) {
    
    # Read the CSV
    temp_data <- read.csv(file, stringsAsFactors = FALSE)
    
    # Skip empty files
    if (nrow(temp_data) == 0) {
      next
    }
    
    # Sample up to 100 rows from this file
    sample_n <- min(100, nrow(temp_data))
    sampled_data <- temp_data[sample(seq_len(nrow(temp_data)), size = sample_n), ]
    
    # Extract filename without .csv
    file_name <- tools::file_path_sans_ext(basename(file))
    
    # Convert filename to Date
    speech_date <- as.Date(file_name, format = "%Y-%m-%d")
    
    # Add metadata columns
    sampled_data$speech_date <- speech_date
    sampled_data$year <- as.integer(format(speech_date, "%Y"))
    sampled_data$month <- as.integer(format(speech_date, "%m"))
    sampled_data$day <- as.integer(format(speech_date, "%d"))
    sampled_data$weekday <- weekdays(speech_date)
    sampled_data$source_file <- basename(file)
    sampled_data$source_path <- file
    
    # Store in list
    sample_list[[counter]] <- sampled_data
    counter <- counter + 1
  }
  
  cat("Finished year", year, "-", i, "of", total_years, "completed\n\n")
}

# Combine all sampled data at the end
master_sample <- do.call(rbind, sample_list)

# Write the sampled master file
write.csv(master_sample,
          file = file.path(master_folder, "master_sample_file.csv"),
          row.names = FALSE)

cat("Done. Sample master file written with", nrow(master_sample),
    "rows and", ncol(master_sample), "columns.\n")
```
This loop creates a master Hansard speech file from the large raw data found 
online from 1900 to 2019. To simualte stratified random sampling, I sampled 100 
rows from each speech file in the dataset, which resulted in 1312018 obs. of 
22 variables, which is still pretty large, so from here, 
I will apply a dictionary-based technique to gather just the education-related speeches. 


##Using Dictionary-Based Aproach to Keep Only Education-Related Speeches
```{r}
#Creating dictionary
education_pattern <- regex(
  "ducation|school|universit|student|tuition|textbook|curriculum|teach|colleg|postsec|post-sec|research|vocational|apprenticeshi|academ",
  ignore_case = TRUE)
#Applying the dictionary
education_speeches <- master_sample %>%
  filter(!is.na(speechtext), speechtext != "", str_detect(speechtext, education_pattern))

write.csv(education_speeches, file = file.path(master_folder, "education_speeches.csv"), row.names = F)
```
Here, I used the regex() function to create a dictionary and separate just the education-related
speeches from my initial sample and apply the dictionary to filter for education-related speeches, which gave me 60,833 obs. of 22 variables. 

##Looking At Trends in Sample
```{r}
educ_over_time <- education_speeches %>% 
  mutate(speechmonth = as.Date(paste0(substr(speechdate, 1, 7), "-01"))) %>% 
  group_by(year) %>% 
  count() %>% 
  mutate(prop = n/60833*100) %>% ungroup()

educ_over_time %>% 
  mean_ci(prop)

ggplot(educ_over_time, aes(x = year, y = prop)) +
  geom_line(color = "orange", linewidth = 1.2) +
  geom_hline(yintercept = .8547009, linetype = "longdash") +
  theme_bw() +
  scale_x_continuous(limits = c(1900, 2020), breaks = seq(1910, 2010, by = 20), expand = c(0,0)) +
  scale_y_continuous(limits = c(0, 2.5), breaks = seq(0.25, 2.25, by = 0.5), expand = c(0,0), labels = function(x) paste0(x, "%")) +
  labs(x = "Year", y = "Sample Proportion of Speeches")
```
Here, I looked at trends in the sample to make sense of any changes in prevalence over time. 
When doing preliminary research, I found that the increases in prevalence 
correspond to periods in Canadian history when education is a salient topic, which makes sense. 
Here, despite the sample proportion of speeches being generally low overall, 
there are still clear spikes which are correlated with national education events. 


##Constructing and Cleaning my Corpus
```{r}
education_speeches <- education_speeches %>% 
  mutate(speechmonth = as.Date(paste0(substr(speechdate, 1, 7), "-01"))) 

#Create a month variable
education_speeches <- education_speeches %>% 
  mutate(clean_text = str_replace_all(speechtext, "\\n", "\\."), 
        clean_text = str_replace_all(clean_text, "\\r", "\\.")) %>% 
  dplyr::select(-speechtext)
education_speeches <- education_speeches %>% mutate(clean_text = str_replace_all(clean_text, "-", ""))
education_speeches <- education_speeches %>% mutate(clean_text = str_replace_all(clean_text, "\\..", "\\."))
education_speeches <- education_speeches %>% mutate(clean_text = str_replace_all(clean_text, "\\;.f", "\\;"))
education_speeches <- education_speeches %>% mutate(clean_text = str_replace_all(clean_text, "\\:.f", "\\:"))
education_speeches <- education_speeches %>% mutate(clean_text = str_replace_all(clean_text, "\\?.", ""))
education_speeches <- education_speeches %>% mutate(clean_text = str_replace_all(clean_text, "\\:.e", "\\:"))
education_speeches$clean_text[100]
```
 Because this corpus is still rather large, I 


##Taking a Sample of the Corpus for Hand-coding
```{r}
obj_sample1<- education_speeches %>% #grabbing speeches from before 1990
  filter(year <= 1990) %>% 
  group_by(speechmonth) %>% 
  slice_sample(n = 6)

obj_sample2<- education_speeches %>% 
  filter(year > 1990) %>% 
  group_by(speechmonth) %>%  #grabbing speeches from after 1990
  slice_sample(n = 12)

sample <- full_join(obj_sample1, obj_sample2)

sample <- sample %>% 
  select(-c(hid, opid, source_file, source_path, speakerurl, pid, speakeroldname, speakerposition, )) #is my corpus

ksample<- sample %>% ungroup() %>% 
  slice_sample(n = 200)

write.csv(ksample,
          file = file.path(master_folder, "khandcoding.csv"),
          row.names = FALSE)
```





##Validating Hand-coding Against LLM coding
```{r}
indig_ed<- read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/zero_shot_indig.csv")
lang_ed<- read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/zero_shot_langed.csv")
afford_ed<- read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/zero_shot_afforded.csv")
univ_ed<- read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/zero_shot_unived.csv")
skills_ed<- read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/zero_shot_skillsed.csv")
elc_ed<- read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/zero_shot_elced.csv")
fed_ed<- read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/zero_shot_feded.csv")

khand <- read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/khandcoding.csv")
indig_ed <- indig_ed[, c(1:16)]

predictions <-full_join(indig_ed, lang_ed)
predictions <-full_join(predictions, afford_ed)
predictions <-full_join(predictions, univ_ed)
predictions <-full_join(predictions, skills_ed)
predictions <-full_join(predictions, elc_ed)
predictions <-full_join(predictions, fed_ed)

newmain <-full_join(predictions, khand)


#Calculating Cohen's Kappa
indigkp <- kappa2(newmain[,c(16,23)], "unweighted") #indig_ed
langedkp <- kappa2(newmain[,c(17,24)], "unweighted") #lang_ed
affordkp <- kappa2(newmain[,c(18,25)], "unweighted") #afford_ed
univkp <- kappa2(newmain[,c(19,26)], "unweighted") #univ_ed
skillskp <- kappa2(newmain[,c(20,27)], "unweighted") #skills_ed
elckp <- kappa2(newmain[,c(21,28)], "unweighted") #elc_ed
fedkp <- kappa2(newmain[,c(22,29)], "unweighted") #fed_ed



#Calculating Krippendorffs Alpha
indigmat <-t(as.matrix(newmain[, c(16, 23)]))
indigka <- kripp.alpha(indigmat, method = "nominal")

langmat <-t(as.matrix(newmain[, c(17, 24)]))
langka <- kripp.alpha(langmat, method = "nominal")

affordmat <-t(as.matrix(newmain[, c(18, 25)]))
affordka <- kripp.alpha(affordmat, method = "nominal")

univmat <-t(as.matrix(newmain[, c(19, 26)]))
univka <- kripp.alpha(univmat, method = "nominal")

skillsmat <-t(as.matrix(newmain[, c(20, 27)]))
skills <- kripp.alpha(skillsmat, method = "nominal")

elcmat <-t(as.matrix(newmain[, c(21, 28)]))
elcka <- kripp.alpha(elcmat, method = "nominal")

fedmat <-t(as.matrix(newmain[, c(22, 29)]))
fedka <- kripp.alpha(fedmat, method = "nominal")


#Making a Table
kappas <- c(indigkp$value, langedkp$value, affordkp$value, univkp$value, skillskp$value, elckp$value, fedkp$value)
alphas <- c(indigka$value, langka$value, affordka$value, univka$value, skills$value, elcka$value, fedka$value)
compar_table <- cbind(kappas, alphas)

print(compar_table)
```
The results of my CK and KA scores are helpful for determining whether my 
current codebook is appropriate for the rest of this project and generally 
testing my expectations of the data. Cohen's alpha scores ranged from .28 to
.66, while Krippendorff's alpha ranged from .275 to .665. The similarty between 
these two measures suggests that there is a clear pattern of agreement, however,
how low these scores are also suggest that the model performance varies 
considerably by label. There were three labels which received moderate scores 
(3, 5, 6), while the other 4 showed weaker agreement. In all, these validation 
results suggest that mentions of education affordability (3), skills/training(5), 
and early learning and childhood education(6) are more reliably identified in 
the data than the other categories. While coding, I did find that education 
was discussed in other ways not covered by my codebook, including more general 
mentions of education as well as immigration/student visas, and religion and 
education. From here, I think it makes sense to update my codebook to account 
for these and do additional fine-tuning after this to try and get more robust findings. 





#Part 2


##Expanding the Training Set/Creating a Dictionary for Each Topic
```{r}

#Creating individual dictionaries for each topic 
ed_indig <- regex(
  "First Nation|Inuit|Métis|Indigenous|residential school|reconciliation|aborigin|native|Indian|treaty nation|treaty right|Charter right",
  ignore_case = TRUE)

ed_lang <- regex(
  "bilingual|multilingual|language|minority|mother tongue|linguist|immersion|oral proficiency|transla|anglophone|francophone|french|english|FSL|ESL|pluraling|OLMC|section 23|fluency",
  ignore_case = TRUE)

ed_afford <- regex(
  "afford|cost|financ||econom|barrier|burden|obstac|price|need-based|aid|expenses|tuition|hike|freeze|increase|fee|childcare|grant|bursar|scholarship|loan|debt|deliquin|arrear|intern|savings|access|RESP|OSAP|student assist|subsid|welfare|mobili|insecur|precar|fellowship",
  ignore_case = TRUE)

ed_univ <- regex(
  "universi|college|campus|facult|department|school|institut|academ|postsec|post-sec|higher ed|board|rector|provost|dean|registrar|admission|enrolment|student|graduat|alumn|tenure|adjunct|professor|lecturer|instructor|scholar|curricul|pedagog|teach|learn|classroom|endowment|research|inquir|investig|discover|experimen|evidence|analysis|methodol|fieldwork|peer review|science|preprint|REB|IRB|interdiscipli|multidiscipli|transdiscipli|postdoc|post-doc|doctoral|dissertat|grant|fund|innovat|prototyp|patent|R&D|development|STEM|science|engineering|medicine|humanities|EDI|retention|DEI|practicum|Tri-Council|SSHRC|NSERC|CIHR|CFI",
  ignore_case = TRUE)

ed_skills <- regex(
  "skill|train|work|employ|job|vocat|technic|apprentic|certific|credent|qualification|competenc|proficien|literacy|numeracy|transferable|practical|occupat|readi|human capital|adult|lifelong|experien|cooperative|co-op|intern|practicum|labour|market|profession|career|hir|recruit|mobili|promotion|advance|produc|trade|craft|applied learn|accred|talent|capaci|compet|indust|business|placement|attain|progression|PLA|counsel|mentor|support|guid",
  ignore_case = TRUE)

ed_elc <- regex(
  "early learn|childcare|child care|early child|ELC|ECE|presch|pre-school|pre-k|kinder|child|daycare|day care|family|parent|infant|toddler|nursery|caregive|matern|patern|CWELCC|ELCC",
  ignore_case = TRUE)

ed_fed <- regex(
  "feder|provinc|territor|municip|local|intergov|multilevel|multi-level|shared gov|divided power|constitution|jurisdic|coordinat|asymmet|region|authority|mandate|regulat|govern|section 23|treaty|Charter|court|judic|litig|statut|concurrent|fragmentation|patchwork|harmoniz|complianc|enforce|implement|spend|expend|fiscal|budget|financ|appropria|line item|allocat|transfer|grant|endow|subsid|tax|performance|minist|cabinet|agency|commission|bureaucra|audit|evaluat|equity|agenda|auster|invest|account",
  ignore_case = TRUE)

#Applying the dictionaries 
indig_speeches <- education_speeches %>%
  filter(str_detect(clean_text, ed_indig))

lang_speeches <- education_speeches %>%
  filter(str_detect(clean_text, ed_lang))

afford_speeches <- education_speeches %>%
  filter(str_detect(clean_text, ed_afford))

univ_speeches <- education_speeches %>%
  filter(str_detect(clean_text, ed_univ))

skills_speeches <- education_speeches %>%
  filter(str_detect(clean_text, ed_skills))

elc_speeches <- education_speeches %>%
  filter(str_detect(clean_text, ed_elc))

fed_speeches <- education_speeches %>%
  filter(str_detect(clean_text, ed_fed))

head(indig_speeches$clean_text)

#Writing Speeches to CSV
write.csv(indig_speeches, 
          file = file.path(master_folder, "indigspeches.csv"), 
          row.names = F)
write.csv(lang_speeches,
          file = file.path(master_folder, "langspeeches.csv"),
          row.names = FALSE)
write.csv(afford_speeches,file = file.path(master_folder, "affordspeeches.csv"),
          row.names = FALSE)
write.csv(univ_speeches,
          file = file.path(master_folder, "univspeeches.csv"),
          row.names = FALSE)
write.csv(skills_speeches,
          file = file.path(master_folder, "skillsspeeches.csv"),
          row.names = FALSE)
write.csv(elc_speeches,
          file = file.path(master_folder, "elcspeeches.csv"),
          row.names = FALSE)
write.csv(fed_speeches,
          file = file.path(master_folder, "fedspeeches.csv"),
          row.names = FALSE)

#Fine Tuning Prep

indig150 <- read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Topic CSVs/indigspeches.csv")
lang150 <- read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Topic CSVs/langspeeches.csv")
afford150 <- read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Topic CSVs/affordspeeches.csv")
univ150 <- read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Topic CSVs/univspeeches.csv")
skills150 <- read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Topic CSVs/skillsspeeches.csv")
elc150 <- read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Topic CSVs/elcspeeches.csv")
fed150 <- read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Topic CSVs/fedspeeches.csv")

#Filtering out the labeled data and unique character
indig150 <- indig150 %>% 
  filter(indg_pos == 1) %>% 
  select(c(basepk, indg_pos))

lang150 <- lang150 %>% 
  filter(langpos == 1) %>% 
  select(c(basepk, langpos))

afford150 <- afford150 %>% 
  filter(affordpos == 1) %>% 
  select(c(basepk, affordpos))

univ150 <- univ150 %>% 
  filter(univpos == 1) %>% 
  select(c(basepk, univpos))

skills150 <- skills150 %>% 
  filter(skillpos == 1) %>% 
  select(c(basepk, skillpos))

elc150 <- elc150 %>% 
  filter(elcpos == 1) %>% 
  select(c(basepk, elcpos))

fed150 <- fed150 %>% 
  filter(fedpos == 1) %>% 
  select(c(basepk, fedpos))

#Joining

indig150  <- indig150  %>% mutate(basepk = as.character(basepk))
lang150   <- lang150   %>% mutate(basepk = as.character(basepk))
afford150 <- afford150 %>% mutate(basepk = as.character(basepk))
univ150   <- univ150   %>% mutate(basepk = as.character(basepk))
skills150 <- skills150 %>% mutate(basepk = as.character(basepk))
elc150    <- elc150    %>% mutate(basepk = as.character(basepk))
fed150    <- fed150    %>% mutate(basepk = as.character(basepk))
newmain   <- newmain   %>% mutate(basepk = as.character(basepk))

positives <-full_join(indig150, lang150)
positives <-full_join(positives,afford150 )
positives <-full_join(positives, skills150)
positives <-full_join(positives, elc150)
positives <-full_join(positives, fed150)
positives <-full_join(positives, univ150)

newmain2 <- full_join(positives, newmain, by = "basepk")

table(newmain2$affordpos)
```

#Create a Function to Make the Training Files
```{r}
khand_full <- read_csv(
  "C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/khandcoding.csv",
  show_col_types = FALSE
) %>%
  mutate(basepk = as.character(basepk))

build_topic_train <- function(topic_file, khand_col, topic_col, out_file) {
  
  topic_df <- read_csv(topic_file, show_col_types = FALSE) %>%
    mutate(basepk = as.character(basepk))
  
  # Original hand-coded sample for this topic
  base_df <- khand_full %>%
  transmute(
    basepk = as.character(basepk),
    clean_text = clean_text,
    label = as.integer(.data[[khand_col]])
  ) %>%
  filter(label %in% c(0, 1))
  
  # Expanded checked topic file for this topic
  extra_df <- topic_df %>%
  transmute(
    basepk = as.character(basepk),
    clean_text = clean_text,
    label = as.integer(.data[[topic_col]])
  ) %>%
  filter(label %in% c(0, 1))
  
  # Combine both sources
  combined <- bind_rows(
    base_df %>% mutate(source = "khand"),
    extra_df %>% mutate(source = "topic")
  )
  
  # Flag duplicates where labels disagree
  conflicts <- combined %>%
    group_by(basepk) %>%
    summarise(
      n_rows = n(),
      n_labels = n_distinct(label),
      labels = paste(sort(unique(label)), collapse = ","),
      .groups = "drop"
    ) %>%
    filter(n_labels > 1)
  
  # Keep one row per basepk; if duplicated, prefer khand version
  final_df <- combined %>%
    arrange(basepk, desc(source == "khand")) %>%
    distinct(basepk, .keep_all = TRUE) %>%
    select(basepk, clean_text, label) %>%
    filter(!is.na(clean_text), clean_text != "")
  
  write_csv(final_df, out_file)
  
  list(
    data = final_df,
    conflicts = conflicts,
    counts = table(final_df$label)
  )
}


#Apply the function for each of the topics and write individual CSVs

indig_train_res <- build_topic_train(
  topic_file = "C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Topic CSVs/indigspeches.csv",
  khand_col = "indig_ed",
  topic_col  = "indg_pos",
  out_file   = "C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/indig_train.csv"
)

indig_train_res$counts


lang_train_res <- build_topic_train(
  topic_file = "C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Topic CSVs/langspeeches.csv",
  khand_col = "lang_ed",
  topic_col  = "langpos",
  out_file   = "C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/lang_train.csv"
)
lang_train_res$counts

afford_train_res <- build_topic_train(
  topic_file = "C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Topic CSVs/affordspeeches.csv",
  khand_col = "afford_ed",
  topic_col  = "affordpos",
  out_file   = "C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/afford_train.csv"
)

afford_train_res$counts

univ_train_res <- build_topic_train(
  topic_file = "C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Topic CSVs/univspeeches.csv",
   khand_col = "univ_ed",
  topic_col  = "univpos",
  out_file   = "C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/univ_train.csv"
)
univ_train_res$counts

skills_train_res <- build_topic_train(
  topic_file = "C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Topic CSVs/skillsspeeches.csv",
   khand_col = "skills_ed",
  topic_col  = "skillpos",
  out_file   = "C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/skills_train.csv"
)
skills_train_res$counts

elc_train_res <- build_topic_train(
  topic_file = "C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Topic CSVs/elcspeeches.csv",
   khand_col = "elc_ed",
  topic_col  = "elcpos",
  out_file   = "C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/elc_train.csv"
)
elc_train_res$counts

fed_train_res <- build_topic_train(
  topic_file = "C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Topic CSVs/fedspeeches.csv",
   khand_col = "fed_ed",
  topic_col  = "fedpos",
  out_file   = "C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/fed_train.csv"
)

fed_train_res$counts

```
#Analysing the Results from the Models and Making Images
```{r}
#reading in the final prediction CVs
in_fin_pred <-read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Final Predictions/indig_predictions.csv")
lan_fin_pred <-read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Final Predictions/lang_predictions.csv")
af_fin_pred <-read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Final Predictions/afford_predictions.csv")
sk_fin_pred <-read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Final Predictions/skills_predictions.csv")
un_fin_pred <-read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Final Predictions/univ_predictions.csv")
el_fin_pred <-read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Final Predictions/elc_predictions.csv")
fe_fin_pred <-read_csv("C:/Users/KaylynnTemp/Documents/Winter2026PhD/Text Analysis/Final Project/Final Predictions/fed_predictions.csv")


final <- full_join(in_fin_pred, lan_fin_pred)
final <- full_join(final, af_fin_pred)
final <- full_join(final, sk_fin_pred)
final <- full_join(final, un_fin_pred)
final <- full_join(final, el_fin_pred)
final <- full_join(final, fe_fin_pred)


final <-final %>%  
  select(c(basepk, speechdate, speechmonth, year, month, day, weekday, 
           speakerparty, clean_text, indig_pred, indig_conf, lang_pred, lang_conf, 
           skills_pred, skills_conf, univ_pred, univ_conf, elc_pred, elc_conf, 
           afford_pred, afford_conf, fed_pred, fed_conf))

final <- final %>% 
  filter(!is.na(indig_pred) & !is.na(lang_pred) & !is.na(skills_pred) & 
           !is.na(univ_pred) & !is.na(elc_pred) & !is.na(afford_pred) & 
           !is.na(fed_pred))



topics <- final %>% 
  group_by(year) %>% 
  summarise(indig = mean(indig_pred),
            lang = mean(lang_pred), 
            afford = mean(afford_pred), 
            skills = mean(skills_pred), 
            univ = mean(univ_pred), 
            elc = mean(elc_pred), 
            fed = mean(fed_pred), 
            n = n(), 
            indig_sd = sd(indig_pred),
            lang_sd = sd(lang_pred), 
            afford_sd = sd(afford_pred), 
            skills_sd = sd(skills_pred), 
            univ_sd = sd(univ_pred), 
            elc_sd = sd(elc_pred), 
            fed_sd = sd(fed_pred)) %>% 
  mutate(indig_upper = indig + 1.96*indig_sd/sqrt(n),
         indig_lower = indig - 1.96*indig_sd/sqrt(n),
         lang_upper = lang + 1.96*lang_sd/sqrt(n),
         lang_lower = lang - 1.96*lang_sd/sqrt(n),
         afford_upper = afford + 1.96*afford_sd/sqrt(n),
         afford_lower = afford - 1.96*afford_sd/sqrt(n),
         skills_upper = skills + 1.96*skills_sd/sqrt(n),
         skills_lower = skills - 1.96*skills_sd/sqrt(n),
         univ_upper = univ + 1.96*univ_sd/sqrt(n),
         univ_lower = univ - 1.96*univ_sd/sqrt(n),
         elc_upper = elc + 1.96*elc_sd/sqrt(n),
         elc_lower = elc - 1.96*elc_sd/sqrt(n),
         fed_upper = fed + 1.96*fed_sd/sqrt(n),
         fed_lower = fed - 1.96*fed_sd/sqrt(n))

topics %>% 
  ungroup() %>% 
  summarize(indig = mean(indig),
            lang = mean(lang), 
            afford = mean(afford), 
            skills = mean(skills), 
            univ = mean(univ), 
            elc = mean(elc), 
            fed = mean(fed))

#Topics By Year
ggplot(topics, aes(x = year)) +
  geom_line(aes(y = indig, ymin = indig_lower, ymax = indig_upper, color = "Indigenous"), linewidth = 1.2) +
   geom_line(aes(y = lang, ymin = lang_lower, ymax = lang_upper, color = "Minority Language"), linewidth = 1.2) +
   geom_line(aes(y = afford, ymin = afford_lower, ymax = afford_upper, color = "Affordability"), linewidth = 1.2) +
   geom_line(aes(y = skills, ymin = skills_lower, ymax = skills_upper, color = "Skill Acquisition"), linewidth = 1.2) +
   geom_line(aes(y = univ, ymin = univ_lower, ymax = univ_upper, color = "Research"),  linewidth = 1.2) +
   geom_line(aes(y = elc, ymin = elc_lower, ymax = elc_upper, color = "Early Learning"),  linewidth = 1.2) +
   geom_line(aes(y = fed, ymin = fed_lower, ymax = fed_upper, color = "Federalism"), linewidth = 1.2) +
  theme_bw() +
  scale_color_manual(values = c("darkgreen", "darkblue", "darkred","black", "darkorange", "darkgoldenrod","darkcyan")) +
  scale_x_continuous(limits = c(1900, 2020), breaks = seq(1910, 2010, by = 20), expand = c(0,0)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0.1, 0.9, by = 0.2), expand = c(0,0), labels = function(x) paste0(x*100, "%")) +
  labs(x = "Year", y = "Sample Proportion of Topics")
```


```{r, fig.width= 15, fig.height= 10}
topicmeans <- final %>% 
  ungroup() %>% 
   summarise(indig_ = mean(indig_pred),
            lang_ = mean(lang_pred), 
            afford_ = mean(afford_pred), 
            skills_ = mean(skills_pred), 
            univ_ = mean(univ_pred), 
            elc_ = mean(elc_pred), 
            fed_ = mean(fed_pred), 
            n = n(), 
            indig_sd = sd(indig_pred),
            lang_sd = sd(lang_pred), 
            afford_sd = sd(afford_pred), 
            skills_sd = sd(skills_pred), 
            univ_sd = sd(univ_pred), 
            elc_sd = sd(elc_pred), 
            fed_sd = sd(fed_pred)) %>% 
  mutate(indig_upper = indig_ + 1.96*indig_sd/sqrt(n),
         indig_lower = indig_ - 1.96*indig_sd/sqrt(n),
         lang_upper = lang_ + 1.96*lang_sd/sqrt(n),
         lang_lower = lang_ - 1.96*lang_sd/sqrt(n),
         afford_upper = afford_ + 1.96*afford_sd/sqrt(n),
         afford_lower = afford_ - 1.96*afford_sd/sqrt(n),
         skills_upper = skills_ + 1.96*skills_sd/sqrt(n),
         skills_lower = skills_ - 1.96*skills_sd/sqrt(n),
         univ_upper = univ_ + 1.96*univ_sd/sqrt(n),
         univ_lower = univ_ - 1.96*univ_sd/sqrt(n),
         elc_upper = elc_ + 1.96*elc_sd/sqrt(n),
         elc_lower = elc_ - 1.96*elc_sd/sqrt(n),
         fed_upper = fed_ + 1.96*fed_sd/sqrt(n),
         fed_lower = fed_ - 1.96*fed_sd/sqrt(n))

temp <- topicmeans %>% 
  pivot_longer(
    cols = matches("^(indig|lang|skills|univ|elc|fed|afford)"), 
    names_to = c("Topic", ".value"), 
    names_pattern = "(indig|lang|skills|univ|elc|fed|afford)(.*)"
  ) %>% 
  mutate(label = case_when(Topic == "indig" ~ "Indigenous", Topic == "lang" ~ "Minority Language", 
                           Topic == "afford" ~ "Affordability", Topic == "skills" ~ "Skill Acquisition", 
                           Topic == "univ" ~ "Research", Topic == "elc" ~ "Early Learning", 
                           Topic == "fed" ~ "Federalism"))


#Overall Proportion of Topic in Sample
ggplot(temp, aes(x = reorder(label, `_`) , y = `_`, ymin = `_lower`, ymax = `_upper`, fill = label)) +
  geom_col(color = "black") +
  geom_errorbar(width = 0.5, color = "black") +
  theme_bw() +
  theme(legend.position = "") +
  scale_fill_manual(values = c("darkgreen", "darkblue", "darkred","black", "darkorange", "darkgoldenrod","darkcyan")) +
  scale_y_continuous(limits = c(0, .8), breaks = seq(0.1, 0.7, by = 0.2), expand = c(0,0), labels = function(x) paste0(x*100, "%")) +
  labs(x = "Education Topic", y = "Proportion of Topics")


#Liberal, Conservative, NDP, New Democratic Party Bloc Québécois, Reform, Social Credit, Progressive Conservative, Conservative (1867-1942), Co-operative Commonwealth Federation (C.C.F.), Canadian Alliance

#Topics by Party
pidtop <- final %>% 
  ungroup() %>% 
  group_by(speakerparty) %>% 
  filter(!is.na(speakerparty)) %>% 
  filter(speakerparty == "Liberal" | speakerparty == "Conservative" | speakerparty == "New Democratic Party" | speakerparty == "Bloc Québécois" | speakerparty == "Reform" |speakerparty == "Social Credit" | speakerparty == "Progressive Conservative" | speakerparty == "Conservative (1867-1942)" | speakerparty == "Co-operative Commonwealth Federation (C.C.F.)" | speakerparty == "Canadian Alliance") %>% 
   summarise(indig_ = mean(indig_pred),
            lang_ = mean(lang_pred), 
            afford_ = mean(afford_pred), 
            skills_ = mean(skills_pred), 
            univ_ = mean(univ_pred), 
            elc_ = mean(elc_pred), 
            fed_ = mean(fed_pred), 
            n = n(), 
            indig_sd = sd(indig_pred),
            lang_sd = sd(lang_pred), 
            afford_sd = sd(afford_pred), 
            skills_sd = sd(skills_pred), 
            univ_sd = sd(univ_pred), 
            elc_sd = sd(elc_pred), 
            fed_sd = sd(fed_pred)) %>% 
  mutate(indig_upper = indig_ + 1.96*indig_sd/sqrt(n),
         indig_lower = indig_ - 1.96*indig_sd/sqrt(n),
         lang_upper = lang_ + 1.96*lang_sd/sqrt(n),
         lang_lower = lang_ - 1.96*lang_sd/sqrt(n),
         afford_upper = afford_ + 1.96*afford_sd/sqrt(n),
         afford_lower = afford_ - 1.96*afford_sd/sqrt(n),
         skills_upper = skills_ + 1.96*skills_sd/sqrt(n),
         skills_lower = skills_ - 1.96*skills_sd/sqrt(n),
         univ_upper = univ_ + 1.96*univ_sd/sqrt(n),
         univ_lower = univ_ - 1.96*univ_sd/sqrt(n),
         elc_upper = elc_ + 1.96*elc_sd/sqrt(n),
         elc_lower = elc_ - 1.96*elc_sd/sqrt(n),
         fed_upper = fed_ + 1.96*fed_sd/sqrt(n),
         fed_lower = fed_ - 1.96*fed_sd/sqrt(n))



pidtemp <- pidtop %>% 
  pivot_longer(
    cols = matches("^(indig|lang|skills|univ|elc|fed|afford)"), 
    names_to = c("Topic", ".value"), 
    names_pattern = "(indig|lang|skills|univ|elc|fed|afford)(.*)"
  ) %>% 
  mutate(label = case_when(Topic == "indig" ~ "Indigenous", Topic == "lang" ~ "Minority Language", 
                           Topic == "afford" ~ "Affordability", Topic == "skills" ~ "Skill Acquisition", 
                           Topic == "univ" ~ "Research", Topic == "elc" ~ "Early Learning", 
                           Topic == "fed" ~ "Federalism"))


#Party Topic Proportion
ggplot(pidtemp, aes(x = reorder(label, `_`) , y = `_`, ymin = `_lower`, ymax = `_upper`, fill = label)) +
  geom_col(color = "black") +
  geom_errorbar(width = 0.5, color = "black") +
  theme_bw() +
  theme(legend.position = "", 
        axis.title = element_text(size = 16), axis.text.x = element_text(size = 10), axis.text.y = element_text(size = 14), strip.text = element_text(size = 14)) +
  facet_wrap(~speakerparty, ncol = 2) +
  scale_fill_manual(values = c("darkgreen", "darkblue", "darkred","black", "darkorange", "darkgoldenrod","darkcyan")) +
  scale_y_continuous(limits = c(0, .9), breaks = seq(0.2, 0.8, by = 0.2), expand = c(0,0), labels = function(x) paste0(x*100, "%")) +
  labs(x = "Education Topic", y = "Proportion of Topics")

ggsave("partyprop.jpg", plot = last_plot(), width = 15, height = 10, dpi = 500)
```
#Exemplar Quotes for Report

```{r}
#Define quotes directly
quote_panels <- c(
  affordability = "Does the minister not realize that besides doubling tuition fees, his reform helps to double students’ indebtedness...?", 
  childcare = "As a result of the early learning and childcare agreement signed with Ontario, Halton will receive 500 new spaces over the next three years. The childcare spaces will primarily be located in or near schools so that junior and senior kindergarten students can benefit from a seamless full day of learning and childcare...",
skills =  "They voted against increased investments in job training. They have voted against every constructive measure: the Canada apprenticeship loan, the Canada job grant, and the incentive grant for new apprentices...")


panel_labels <- c("A", "B", "C")

jpeg("figure_exemplar_quotes.jpg", width = 10, height = 3.5, units = "in", res = 300)

# smaller margins so panels sit closer together
par(mfrow = c(1, 3),
    mar = c(0.3, 0.3, 1.2, 0.3),   # inner margins around each panel
    oma = c(0, 0, 1.5, 0),         # outer margins for overall title
    xpd = NA)

for (i in seq_along(quote_panels)) {
  
  plot.new()
  plot.window(xlim = c(0, 1), ylim = c(0, 1))
  
  # smaller box inside each panel
  rect(0.08, 0.12, 0.92, 0.88, border = "grey60", lwd = 1.5)
  
  # panel label
  text(
    x = 0.12, y = 0.90,
    labels = panel_labels[i],
    cex = 0.9,
    font = 2,
    family = "sans"
  )
  
  # topic in uppercase
  text(
    x = 0.5, y = 0.78,
    labels = toupper(names(quote_panels)[i]),
    cex = 0.85,
    font = 2,
    family = "sans"
  )
  
  # dividing line under topic
  segments(0.20, 0.70, 0.80, 0.70, lwd = 1.1, col = "grey50")
  
  # wrapped quote
  quote_text <- paste0(
    "“",
    paste(strwrap(quote_panels[i], width = 24), collapse = "\n"),
    "”"
  )
  
  text(
    x = 0.5, y = 0.42,
    labels = quote_text,
    cex = 0.88,
    family = "sans"
  )
}

mtext("Exemplar Hansard Quotes",
      outer = TRUE, cex = 1.0, font = 2)

dev.off()
```

